Implementing a Capability Machine model into Iris

### 2020-01-11 prisc-2020-talk/slides.tex
moved the execute condition and expression relation to bonus slides

%------------------------------------------------
\subsection{The Execute Condition}

@@

 	\begin{align*}
 		\mathcal{E}\interp{W}{pc} \triangleq&~\forall r, \mathcal{R}(W)(r) ~*~{\only<2->{\color{red}} \textsf{context}(W)(r[\textsf{\small PC}:=pc])}\\
 		&\sep~\textsc{WP}~\textsf{Seq (Instr Executable)}~\\
 		&\hspace{1cm}\{ v, v = HaltedV \implies \exists W' r', W' \sqsubseteq_{priv} W \\
 		&\hspace{1cm}* {\only<2->{\color{red}}\textsf{context}(W')(r')}\}
 	\end{align*}

\only<2->{
\begin{align*}
	{\color{red} \textsf{context}(W)(r) =~\only<2>{?}} &\only<3->{(\bigsep_{r_i \mapsto w \in r} r_i \mapsto_r w) \wedge \textsf{full\_map } r}\\
	\only<4->{&* \textsf{na\_inv } \gamma_{na} \top \\ }
	\only<5->{&* \textsf{sts\_full } W \\ }
	\only<6->{&* \textsf{region } W \\ }
\end{align*}
}

\begin{frame}
\begin{itemize}
	\item<2-> "If we can read a region" : $p = \textsc{\small RX} \vee p = \textsc{\small RWX} \vee p = \textsc{\small RWLX}$
	\item<3-> "and every word in that region is safe": $\textsf{read\_write\_cond }(p,b,e)$
	\item<4-> "then we can safely execute it": $\mathcal{E}\interp{W}{((p,g),b,e,a)}$
\end{itemize}

\begin{align*}
	&(p = \textsc{\small RX} \vee p = \textsc{\small RWX} \vee p = \textsc{\small RWLX}) \implies\\
	&\textsf{read\_write\_cond }(p,b,e) \implies \mathcal{E}\interp{W}{((p,g),b,e,a)}
\end{align*}
\end{frame}

\begin{align*}
 		\mathcal{E}\interp{W}{pc} \triangleq&~\forall r, \mathcal{R}(W)(r) ~*~{\textsf{context}(W)(r[\textsf{\small PC}:=pc])}\\
 		&\sep~\textsc{WP}~\textsf{Seq (Instr Executable)}~\\
 		&\hspace{1cm}\{ v, v = HaltedV \implies \exists W' r', W' \sqsubseteq_{priv} W \\
 		&\hspace{1cm}* {\textsf{context}(W')(r')}\}
 	\end{align*}

@@

\begin{itemize}
	\item Mechanized formalization: currently $\sim 20,000$ lines of code
	\item A higher level of abstraction
\end{itemize}

~\\[1em]
\textbf{What's new: Iris formalization viewpoint}
\begin{itemize}
	\item Formalization of a machine level language, with no distinction between program and memory
	\item Distinction between  well-bracketed and non well-bracketed calls: using public/private transitions
\end{itemize}

@@

%----------------------------------------------------------------------------------------

>>>>

\begin{frame}
\frametitle{From World to state transition system collection}
On paper: 
\begin{align*}
	\text{Region} =~&\{Revoked\}~\uplus \\
						     &\{Temporary\} \times {\color{red} \text{State} \times \text{Rels}} \\ 
						     &\times (\text{State} \rightarrow (Wor \xrightarrow[\sqsupseteq_{pub}]{mon, ne} \text{UPred(MemSeg)}))~\uplus \\
						     &\{Permanent\} \times {\color{red} \text{State} \times \text{Rels}}  \\ 
						     &\times (\text{State} \rightarrow (Wor \xrightarrow[\sqsupseteq_{priv}]{mon, ne} \text{UPred(MemSeg)})) \\
	\text{World} =~&\mathds{N} \rightharpoonup \text{Region}
\end{align*}
In the Iris mechanization, we use a collection of state transition systems: 
$$\Sigma: \mathds{N} \rightharpoonup States \times \mathds{N} \rightharpoonup Rels $$
\uncover<2->{The world circularity problem is now handled using Iris invariants and saved predicates.}

@@

\begin{frame}
\frametitle{Standard STS}

$$\Sigma: \mathds{N} \rightharpoonup States \times \mathds{N} \rightharpoonup Rels $$

\begin{center}
\begin{tikzpicture}[scale=0.2]
\tikzstyle{every node}+=[inner sep=0pt]
\draw [black] (17.8,-20.1) circle (3);
\draw (17.8,-20.1) node {$temp$};
\draw [black] (39.8,-20.1) circle (3);
\draw (39.8,-20.1) node {$rev$};
\draw [black] (29.2,-33) circle (3);
\draw (29.2,-33) node {$perm$};
\draw [black,dashed] (20.396,-18.602) arc (115.58004:64.41996:19.465);
\fill [black] (37.2,-18.6) -- (36.7,-17.81) -- (36.27,-18.71);
\draw [black] (37.061,-21.318) arc (-69.63899:-110.36101:23.742);
\fill [black] (20.54,-21.32) -- (21.12,-22.07) -- (21.46,-21.13);
\draw [black,dashed] (19.79,-22.35) -- (27.21,-30.75);
\fill [black] (27.21,-30.75) -- (27.06,-29.82) -- (26.31,-30.48);
\draw [black,dashed] (37.9,-22.42) -- (31.1,-30.68);
\fill [black] (31.1,-30.68) -- (32,-30.38) -- (31.23,-29.75);
\draw [black] (16.477,-17.42) arc (234:-54:2.25);
\fill [black] (19.12,-17.42) -- (20,-17.07) -- (19.19,-16.48);
\draw [black] (38.477,-17.42) arc (234:-54:2.25);
\fill [black] (41.12,-17.42) -- (42,-17.07) -- (41.19,-16.48);
\draw [black] (30.523,-35.68) arc (54:-234:2.25);
\fill [black] (27.88,-35.68) -- (27,-36.03) -- (27.81,-36.62);
\end{tikzpicture}
\end{center}

@@

\begin{itemize}
	\item Mechanized formalization: currently $\sim 25 000$ lines of Iris code
	\item At a higher level of abstraction
	\begin{itemize}
		\item<2-> Step index $\rightarrow$ later modality
		\item<3-> World $\rightarrow$ collection of state transition systems
	\end{itemize}
\end{itemize}

~\\[0.5em]
\textbf{What's new: Iris formalization viewpoint}
\begin{itemize}
	\item Formalization of a machine level language, with no distinction between program and memory
	\item Distinction between  well-bracketed and non well-bracketed calls: using public/private transitions
\end{itemize}

@@

%----------------------------------------------------------------------------------------
% Bonus Slides
%------------------------------------------------
\subsection{The Execute Condition}

 	\begin{align*}
 		\mathcal{E}\interp{\Sigma}{pc} \triangleq&~\forall r, \mathcal{R}(\Sigma)(r) ~*~{\only<2->{\color{red}} \textsf{context}(\Sigma)(r[\textsf{\small PC}:=pc])}\\
 		&\sep~\textsc{WP}~\textsf{Seq (Instr Executable)}~\\
 		&\hspace{1cm}\{ v, v = HaltedV \implies \exists \Sigma ' r', \Sigma ' \sqsupseteq_{priv} \Sigma \\
 		&\hspace{1cm}* {\only<2->{\color{red}}\textsf{context}(\Sigma ')(r')}\}
 	\end{align*}

\only<2->{
\begin{align*}
	{\color{red} \textsf{context}(\Sigma)(r) =~\only<2>{?}} &\only<3->{(\bigsep_{r_i \mapsto w \in r} r_i \mapsto_r w) \wedge \textsf{full\_map } r}\\
	\only<4->{&* \textsf{na\_inv } \gamma_{na} \top \\ }
	\only<5->{&* \textsf{sts\_full } \Sigma \\ }
	\only<6->{&* \textsf{region } \Sigma \\ }
\end{align*}
}

\begin{frame}
\begin{itemize}
	\item<2-> "If we can read a region" : $p = \textsc{\small RX} \vee p = \textsc{\small RWX} \vee p = \textsc{\small RWLX}$
	\item<3-> "and every word in that region is safe": $\textsf{read\_write\_cond }(p,b,e)$
	\item<4-> "then we can safely execute it": $\mathcal{E}\interp{\Sigma}{((p,g),b,e,a)}$
\end{itemize}

\begin{align*}
	&(p = \textsc{\small RX} \vee p = \textsc{\small RWX} \vee p = \textsc{\small RWLX}) \implies\\
	&\textsf{read\_write\_cond }(p,b,e) \implies \mathcal{E}\interp{\Sigma}{((p,g),b,e,a)}
\end{align*}
\end{frame}

\begin{align*}
 		\mathcal{E}\interp{\Sigma}{pc} \triangleq&~\forall r, \mathcal{R}(\Sigma)(r) ~*~{\textsf{context}(\Sigma)(r[\textsf{\small PC}:=pc])}\\
 		&\sep~\textsc{WP}~\textsf{Seq (Instr Executable)}~\\
 		&\hspace{1cm}\{ v, v = HaltedV \implies \exists \Sigma ' r', \Sigma ' \sqsupseteq_{priv} \Sigma \\
 		&\hspace{1cm}* {\textsf{context}(\Sigma ')(r')}\}
 	\end{align*}

\begin{frame}
\frametitle{Proving the Fundamental Theorem}
